### 2025-03-14 paper/quarto_ecml/paper.tex
actionability done

In panel (c), we have used conventional training again, this time
imposing the mutability constraint on \emph{age} at test time.
Counterfactuals are valid but involve some substantial reductions in
\emph{debt} for some individuals (very young applicants). By comparison,
counterfactual paths are shorter on average in panel (d), where we have
used CT and protected the immutable feature as described in
Section~\ref{sec-constraints}. We observe that due to the classifier's
lower sensitivity to \emph{age}, recourse recommendations with respect
to \emph{debt} are much more homogenous and do not disproportionately
punish younger individuals. The counterfactuals are also plausible with
respect to the mutable feature. Thus, we consider the model in panel (d)
as the most explainable according to Def. \ref{def-explainability}.

@@

\subsubsection{Plausibility}\label{plausibility}

Table~\ref{tbl-main} presents our main empirical findings. The top five
rows show the percentage reduction in implausibility according to
Equation~\ref{eq-impl-dist} for varying degrees of the energy penalty
used for \emph{ECCo} at test time. The following row shows the reduction
in implausibility as measured by Equation~\ref{eq-impl-div} and
aggregated across all test specifications of \emph{ECCo}. The final two
rows show the test accuracies for the model trained with CT and
conventionally trained models (``vanilla'').

@@

\caption{\label{tbl-main}Key plausibility and predictive performance
metrics for all datasets. The top five rows show the percentage
reduction in implausibility according to Equation~\ref{eq-impl-dist} for
varying degrees of the energy penalty used for \emph{ECCo} at test time.
The following row shows the reduction in implausibility as measured by
Equation~\ref{eq-impl-div} and aggregated across all test specifications
of \emph{ECCo}. The final two rows show the test accuracies for the
model trained with CT and conventionally trained models (``vanilla'').}

@@

\subsubsection{Actionability}\label{actionability}

>>>>

In panel (c), we have used conventional training again, this time
imposing the mutability constraint on \emph{age} at test time.
Counterfactuals are valid but involve some substantial reductions in
\emph{debt} for some individuals (very young applicants). By comparison,
counterfactual paths are shorter on average in panel (d), where we have
used CT and protected the immutable feature as described in
Section~\ref{sec-constraints}. We observe that due to the classifier's
lower sensitivity to \emph{age}, recourse recommendations with respect
to \emph{debt} are much more homogenous and do not disproportionately
punish younger individuals. The counterfactuals are also plausible with
respect to the mutable feature and shorter on average than in (c). Thus,
we consider the model in panel (d) as the most explainable according to
Def. \ref{def-explainability}.

@@

\subsubsection{Plausibility}\label{sec-plaus}

Table~\ref{tbl-main} presents our main empirical findings. The top five
rows show the percentage reduction in implausibility according to
Equation~\ref{eq-impl-dist} for varying degrees of the energy penalty
used for \emph{ECCo} at test time. The following row shows the reduction
in implausibility as measured by Equation~\ref{eq-impl-div} and
aggregated across all test specifications of \emph{ECCo}. The following
two rows show the test accuracies for the model trained with CT and
conventionally trained models (``vanilla'').

@@

\caption{\label{tbl-main}Key plausibility and predictive performance
metrics for all datasets. The top five rows show the percentage
reduction in implausibility according to Equation~\ref{eq-impl-dist} for
varying degrees of the energy penalty used for \emph{ECCo} at test time.
The following row shows the reduction in implausibility as measured by
Equation~\ref{eq-impl-div} and aggregated across all test specifications
of \emph{ECCo}. The following two rows show the test accuracies for the
model trained with CT and conventionally trained models (``vanilla'').
The final row present the average reduction in costs when imposing
mutability constraints (Section~\ref{sec-act})}

@@

\subsubsection{Actionability}\label{sec-act}

In Section~\ref{sec-method}, we show that our proposed way for encoding
mutability constraints leads to lower classifier sensitivity wrt.
immutable features for linear models, tilting the decision boundary in
favour of mutable features instead. For binding constraints at test
time, this has the effect of shorter counterfactual paths and hence
smaller average costs to individuals. To extend this to the non-linear
case, we test the effect of imposing mutability constraints empirically
for our synthetic data using the same evaluation scheme as above. The
final row in Table~\ref{tbl-main} reports the average reduction in costs
for CT compared to the baseline, when imposing that either the first or
the second feature is immutable. In all cases, costs are reduced
substantially indicating that classifiers trained with CT are indeed
more sensitive to mutable features.

### 2025-03-14 paper/sections/experiments.qmd
actionability done

### Plausibility

@tbl-main presents our main empirical findings. The top five rows show the percentage reduction in implausibility according to @eq-impl-dist for varying degrees of the energy penalty used for *ECCo* at test time. The following row shows the reduction in implausibility as measured by @eq-impl-div and aggregated across all test specifications of *ECCo*. The final two rows show the test accuracies for the model trained with CT and conventionally trained models ("vanilla").

@@

Key plausibility and predictive performance metrics for all datasets. The top five rows show the percentage reduction in implausibility according to @eq-impl-dist for varying degrees of the energy penalty used for *ECCo* at test time. The following row shows the reduction in implausibility as measured by @eq-impl-div and aggregated across all test specifications of *ECCo*. The final two rows show the test accuracies for the model trained with CT and conventionally trained models ("vanilla").

@@

### Actionability

>>>>

### Plausibility {#sec-plaus}

@tbl-main presents our main empirical findings. The top five rows show the percentage reduction in implausibility according to @eq-impl-dist for varying degrees of the energy penalty used for *ECCo* at test time. The following row shows the reduction in implausibility as measured by @eq-impl-div and aggregated across all test specifications of *ECCo*. The following two rows show the test accuracies for the model trained with CT and conventionally trained models ("vanilla").

@@

Key plausibility and predictive performance metrics for all datasets. The top five rows show the percentage reduction in implausibility according to @eq-impl-dist for varying degrees of the energy penalty used for *ECCo* at test time. The following row shows the reduction in implausibility as measured by @eq-impl-div and aggregated across all test specifications of *ECCo*. The following two rows show the test accuracies for the model trained with CT and conventionally trained models ("vanilla"). The final row present the average reduction in costs when imposing mutability constraints (@sec-act)

@@

### Actionability {#sec-act}

In @sec-method, we show that our proposed way for encoding mutability constraints leads to lower classifier sensitivity wrt. immutable features for linear models, tilting the decision boundary in favour of mutable features instead. For binding constraints at test time, this has the effect of shorter counterfactual paths and hence smaller average costs to individuals. To extend this to the non-linear case, we test the effect of imposing mutability constraints empirically for our synthetic data using the same evaluation scheme as above. The final row in @tbl-main reports the average reduction in costs for CT compared to the baseline, when imposing that either the first or the second feature is immutable. In all cases, costs are reduced substantially indicating that classifiers trained with CT are indeed more sensitive to mutable features.
